Synthesizing Symmetric Lenses

### 2018-07-08 papers/SymmetricLenses/main.tex
add lotta stuff

\subsection{Simple Symmetric Lenses}
A simple symmetric lens between formats described by types $\BRegex$
and $\RegexAlt$ 
is defined by four functions\bcp{I think a picture would be helpful.  I can probably
  make one pretty easily by tweaking the one from the original symmetric paper.}
\afm{That would be awesome.}

@@

\begin{gather}
 \tag{1}
 \mbox{$\CreateR \OfType \Regex \rightarrow \RegexAlt$}\\
 \tag{2}
 \mbox{$\CreateL \OfType \RegexAlt \rightarrow \Regex$}\\
 \tag{3}
 \mbox{$\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$}\\
 \tag{4}
 \mbox{$\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$}
\end{gather}

@@

We use regular expressions to denote the types of data formats, and we
impose a typing discipline on our lenses.  Specifically, if 
$\Regex$ and $\RegexAlt$ are regular expressions, then the notation
\lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$} indicates that $\Lens$
is a simple symmetric lens between $\LanguageOf{\Regex}$ and
$\LanguageOf{\RegexAlt}$. 
When clear from context, we use the term \textit{lens} to denote a
simple symmetric lens.

@@

The simplest combinator is the identity lens \IdentityLens, which takes as an
argument a regular expression $\Regex$ and performs an identity mapping on data
matching $\Regex$. The identity lens fully connects the
corresponding data in the two formats: when one format is
updated, the other format gets the exact same value. In our example,
when a \lstinline{name} is updated in one file, the identify lens ensures it is
also updated in the other file. In this manner, both the \CreateR and \CreateL
functions are the identity function ($\CreateR \App s = s$), and the put
functions merely return the first argument ($\PutROf{s_1}{s_2} = s_2$).

@@

A challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disconnect(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is another valid lens (where \lstinline{rep($\Regex$)} is a string in the
language of $\Regex$). Furthermore, adding the input-output example in
Figure~\ref{fig:minimized-representations} 
helps, but only a little, because it requires many examples to rule
out all the possible occurrences of nested \Disconnect lenses,
particularly in complex formats. 
%\ksf{I don't understand the following.  We shouldn't explain the issue in
%terms of \PutR and \PutL as we are focused on the combinators rather
%than the raw functions.}
%The default strings given to \Disconnect merely need to be the strings
%in the input-output examples, and the lens will still satisfy the
%specification.  While the disconnect lens is prohibited by two
%examples, or by a carefully chosen input-output example on \PutR
%or \PutL, this problem continues presenting itself for the sublenses,
%and is highlighted even more on larger formats.
While we could ask programmers to provide detailed and precise specifications
for their synthesis tasks, we would prefer an algorithm that can intelligently
choose a lens despite the underspecification. Instead of merely finding
\emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
likely to please the user.

@@

In our setting, we already have a way of expressing sets of data:
regular expressions.  What we need is a way to express probability
distributions over that data.
To that end, we adopt \emph{stochastic regular expressions}\cite{?}, 
which are regular expressions where each operator is annotated with a
probability.  Stochastic regular expressions
allow us to jointly specify the allowable data and give a probability
distribution over it.  From this information, we can calculate the
entropy of the stochastic regular expression.

This infrastructure gives us what we need to find likely
satisfying lenses from a collection of satisfying lenses.  All such
lenses must synchronize data between the same two 
stochastic regular expressions $\Regex$ and $\RegexAlt$.
For any such lens $\Lens$, we can calculate the expected number of
bits required 
to recover a string $s$ in $\LanguageOf{\Regex}$ from a synchronized
string $t$ in $\LanguageOf{\RegexAlt}$. 
This expected number of bits is just the conditional entropy of
$\Regex$ given $\RegexAlt$ and $\Lens$. 
To ensure symmetry, the likelihood we assign to $\Lens$ is the sum of 
the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and
the conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$.

To obtain stochastic regular expressions from the plain regular expressions that
we expect users to supply, we use a heurstic that attempts to assign probability
annotations that give each string in the language equal probablity. We can also
take as input a stochastic regular expression rather than a plain one, which
allows the user to specify a data-specific probability distribution.
Alternatively, the user could provide a large suite of example data from which
we could infer the necessary annotations to model the supplied data. We found
our heuristically chosen distribution works well in practice.

@@

\section{Stochastic Regular Expressions}
\label{sec:srx}
To characterize the likely lenses, we must compute the expected number of bits
needed to recover a string in one data source from a synchronized string in the
other data source. To do this, we must first develop a probabilistic langauge
model for our language. We do this with \emph{stochastic regular expressions},
regular expressions annotated with probability information. With this, we can
jointly express a language and a probability distribution over that language.

@@

In the stochastic regular expression
$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
represents the likelihood the string is in $\Regex_1$. In the stochastic regular
expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability how
much more likely having a string involving more than $n$ iterations is than
having a string involving $n$ iterations. The semantics of SREs is defined by
the function $P$ below.

@@

% begin ssl
\section{Simple Symmetric String Lenses}
\label{sec:ssl}
We now give a formal description of all the combinators in our symmetric lens
language. These lens terms are typed between pairs of stochastic regular
expressions: the typing judgment $\Lens \OfType \Regex \Leftrightarrow
\RegexAlt$ means that $\Lens$ is a well typed lens between $\LanguageOf{\Regex}$
and $\LanguageOf{\RegexAlt}$. These lens terms can also be typed between regular
expressions, but we present in terms of stochastic regular expressions to assist
in describing the entropy of the lenses.

\begin{center}
  \begin{tabular}{l@{\ }l@{\ }l@{\ }l}
    % REGEX
    \Lens{} & \GEq{} & & $\IdentityLensOf{\Regex}$\\
            & & \GBar{} & $\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$ \\
            & & \GBar{} & $\IterateLensOf{\Lens}$ \\
            & & \GBar{} & $\ConcatLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\SwapLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\OrLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\ComposeLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\MergeROf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\MergeLOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\InvertOf{\Lens}$ \\
  \end{tabular}
\end{center}

@@

\[
  \centering
  \inferrule*
  {
  }
  {
    \IdentityLensOf{\Regex} \OfType \Regex \Leftrightarrow \Regex
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App s$ & = & $s$\\
    $\CreateL{} \App s$ & = & $s$\\
    $\PutR{} \App s_1 \App s_2$ & = & $s_1$\\
    $\PutL{} \App s_1 \App s_2$ & = & $s_1$
  \end{tabular}
\end{center}

@@

\[
  \centering
  \inferrule*
  {
    \String \in \LanguageOf{\Regex}\\
    \StringAlt \in \LanguageOf{\RegexAlt}
  }
  {
    \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}
    \OfType \Regex \Leftrightarrow \RegexAlt
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String'$ & = & $\StringAlt$\\
    $\CreateL{} \App \StringAlt'''''$ & = & $\String$\\
    $\PutR{} \App \String''' \App \StringAlt'$ & = & $\StringAlt'$\\
    $\PutL{} \App \StringAlt' \App \String'$ & = & $\String'$
  \end{tabular}
\end{center}
Just as the identity lens ignores the second argument in puts, disconnect lenses
ignore the first in both puts and creates.  The data is unsynchronized in these
two formats, information from one format doesn't impact the other.

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
  }
  {
    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Concat \Regex_2
    \Leftrightarrow
    \RegexAlt_1 \Concat \RegexAlt_2
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1 \String_2$ & = & $(\Lens_1.\CreateROf{\String_1})(\Lens_2.\CreateROf{\String_2})$\\
%    $\CreateL{} \App \StringAlt_1 \StringAlt_2$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
%    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_1\StringAlt_2)$ & = & $(\Lens_1.\PutROf{\String_1}{\StringAlt_1})(\Lens_2.\PutROf{\String_2}{\StringAlt_2})$\\
%    $\PutL{} \App (\StringAlt_1\StringAlt_2) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
%  \end{tabular}
%\end{center}

@@

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
  }
  {
    \SwapLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Concat \Regex_2
    \Leftrightarrow
    \RegexAlt_2 \Concat \RegexAlt_1
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1\String_2$ & = & $(\Lens_2.\CreateROf{\String_2})(\Lens_1.\CreateROf{\String_1})$\\
%    $\CreateL{} \App \StringAlt_2\StringAlt_1$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
%    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_2\StringAlt_1)$ & = & $(\Lens_2.\PutROf{\String_2}{\StringAlt_2})(\Lens_1.\PutROf{\String_1}{\StringAlt_1})$\\
%    $\PutL{} \App (\StringAlt_2\StringAlt_1) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
%  \end{tabular}
%\end{center}
The swap combinator is similar to concat, though the second regular expression
is swapped.

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
  }
  {
    \OrLensOf{\Lens_1}{\Lens_2} \OfType
    \RegexOr{\Regex_1}{\Regex_2}
    \Leftrightarrow
    \RegexOr{\RegexAlt_1}{\RegexAlt_2}
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String$
    & =
    & $\begin{cases*}
      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_1}$\\
      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_2}$
      \end{cases*}$\\

    $\CreateL{} \App \StringAlt$
    & =
    & $\begin{cases*}
      \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1}$\\
      \Lens_2.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_2}$
      \end{cases*}$\\

    $\PutR{} \App \String \App \StringAlt$
    & =
    & $\begin{cases*}
        \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\Regex_1} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_1}$\\
        \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\Regex_2} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_2}$\\
        \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_1} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_2}$\\
        \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_2} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_1}$
      \end{cases*}$\\

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_1} \BooleanAnd \String\in\LanguageOf{\Regex_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2} \BooleanAnd \String\in\LanguageOf{\Regex_2}$\\
        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1} \BooleanAnd \String\in\LanguageOf{\Regex_2}$\\
        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2} \BooleanAnd \String\in\LanguageOf{\Regex_1}$
      \end{cases*}$\\
  \end{tabular}
\end{center}
The \OrLens lens deals with data that can come in one form or another. If the
data gets changed from one format to the other, information in the old format is
lost.

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
  }
  {
    \IterateLensOf{\Lens} \OfType
    \StarOf{\Regex}
    \Leftrightarrow
    \StarOf{\RegexAlt}
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1\ldots\String_n$
%    & =
%    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
%    
%    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
%    & =
%    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
%    
%    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
%    & =
%    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
%        \Lens.\CreateROf{\String_i} & otherwise
%      \end{cases*}$\\
%    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
%    & =
%    & $\String_1'\ldots\String_n'$ where $\String_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
%        \Lens.\CreateROf{\StringAlt_i} & otherwise
%      \end{cases*}$
%  \end{tabular}
%\end{center}
The \IterateLens lens deals with iterated data.

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt
  }
  {
    \MergeROf{\Lens_1}{\Lens_2} \OfType
    \RegexOr{\Regex_1}{\Regex_2}
    \Leftrightarrow
    \RegexAlt
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String$
    & =
    & $\begin{cases*}
      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_1}$\\
      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_2}$
      \end{cases*}$\\

>>>>

\subsection{Simple Symmetric Lenses}
A simple symmetric lens between formats described by regular expression types
$\BRegex$ and $\BRegexAlt$\footnote{We use nonterminals \BRegex and \BRegexAlt
  for regular expressions, using \Regex and \RegexAlt for stochastic regular
  expressions.} is defined by four functions.\bcp{I think a picture would be
  helpful. I can probably make one pretty easily by tweaking the one from the
  original symmetric paper.} \afm{That would be awesome.}

@@

\begin{gather}
 \tag{1}
 \mbox{$\CreateR \OfType \BRegex \rightarrow \BRegexAlt$}\\
 \tag{2}
 \mbox{$\CreateL \OfType \BRegexAlt \rightarrow \BRegex$}\\
 \tag{3}
 \mbox{$\PutR \OfType \BRegex \rightarrow \BRegexAlt \rightarrow \BRegexAlt$}\\
 \tag{4}
 \mbox{$\PutL \OfType \BRegexAlt \rightarrow \BRegex \rightarrow \BRegex$}
\end{gather}

@@

We use regular expressions to denote the types of data formats, and we
impose a typing discipline on our lenses.  Specifically, if 
$\BRegex$ and $\BRegexAlt$ are regular expressions, then the notation
\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that $\Lens$
is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
$\LanguageOf{\BRegexAlt}$. 
When clear from context, we use the term \textit{lens} to denote a
simple symmetric lens.

@@

The simplest combinator is the identity lens \IdentityLens, which takes as an
argument a regular expression $\BRegex$ and performs an identity mapping on data
matching $\BRegex$. The identity lens fully connects the
corresponding data in the two formats: when one format is
updated, the other format gets the exact same value. In our example,
when a \lstinline{name} is updated in one file, the identify lens ensures it is
also updated in the other file. In this manner, both the \CreateR and \CreateL
functions are the identity function ($\CreateR \App s = s$), and the put
functions merely return the first argument ($\PutROf{s_1}{s_2} = s_2$).

@@

A challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disconnect(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is another valid lens (where \lstinline{rep($\BRegex$)} is a string in the
language of $\BRegex$). Furthermore, adding the input-output example in
Figure~\ref{fig:minimized-representations} 
helps, but only a little, because it requires many examples to rule
out all the possible occurrences of nested \Disconnect lenses,
particularly in complex formats. 
%\ksf{I don't understand the following.  We shouldn't explain the issue in
%terms of \PutR and \PutL as we are focused on the combinators rather
%than the raw functions.}
%The default strings given to \Disconnect merely need to be the strings
%in the input-output examples, and the lens will still satisfy the
%specification.  While the disconnect lens is prohibited by two
%examples, or by a carefully chosen input-output example on \PutR
%or \PutL, this problem continues presenting itself for the sublenses,
%and is highlighted even more on larger formats.
While we could ask programmers to provide detailed and precise specifications
for their synthesis tasks, we would prefer an algorithm that can intelligently
choose a lens despite the underspecification. Instead of merely finding
\emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
likely to please the user.

@@

In our setting, we already have a way of expressing sets of data:
regular expressions.  What we need is a way to express probability
distributions over that data.
To that end, we adopt \emph{stochastic regular expressions}\cite{?} (SREs), 
which are regular expressions where each operator is annotated with a
probability.  Stochastic regular expressions
allow us to jointly specify the allowable data and give a probability
distribution over it.  From this information, we can calculate the
entropy of the SRE.

This infrastructure gives us what we need to find likely satisfying lenses from
a collection of satisfying lenses. All such lenses must synchronize data between
the same two SREs $\Regex$ and $\RegexAlt$. For any such lens $\Lens$, we can
calculate the expected number of bits required to recover a string $s$ in
$\LanguageOf{\Regex}$ from a synchronized string $t$ in
$\LanguageOf{\RegexAlt}$. This expected number of bits is just the conditional
entropy of $\Regex$ given $\RegexAlt$ and $\Lens$. To ensure symmetry, the
likelihood we assign to $\Lens$ is the sum of the conditional entropy of
$\Regex$ given $\RegexAlt$ and $\Lens$ and the conditional entropy of
$\RegexAlt$ given $\Regex$ and $\Lens$.

To obtain SREs from the plain regular expressions that
we expect users to supply, we use a heurstic that attempts to assign probability
annotations that give each string in the language equal probablity. We can also
take as input a stochastic regular expression rather than a plain one, which
allows the user to specify a data-specific probability distribution.
Alternatively, the user could provide a large suite of example data from which
we could infer the necessary annotations to model the supplied data. We found
our heuristically chosen distribution works well in practice.

@@

\section{Stochastic Regular Expressions}
\label{sec:srx}
To characterize the likely lenses, we must compute the expected number of bits
needed to recover a string in one data source from a synchronized string in the
other data source. To do this, we must first develop a probabilistic langauge
model for our language. We do this with \emph{stochastic regular expressions},
regular expressions annotated with probability information. With this, we can
jointly express a language and a probability distribution over that language. In
the following grammar, $\String$ ranges over strings, and $\Probability$ ranges
over real numbers between 0 and 1, exclusive.

@@

In the stochastic regular expression
$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation $\Probability$
represents the likelihood the string is in $\Regex_1$. In the stochastic regular
expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability how
much more likely having a string involving more than $n$ iterations is than
having a string involving $n$ iterations. The semantics of SREs is defined by
the function $P$ below.

@@

% begin ssl
\section{Simple Symmetric String Lenses}
\label{sec:ssl}
We now give a formal description of all the combinators in our symmetric lens
language. These lens terms are typed between pairs of stochastic regular
expressions: the typing judgment $\Lens \OfType \BRegex \Leftrightarrow
\BRegexAlt$ means that $\Lens$ is a well typed lens between
$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$. While we are performing
synthesis on stochastic regular expressions, lenses are typed between regular
expressions.  We can recover a regular expression from a stochastic regular
expression by deleting probability annotations.  In a slight abuse of notation,
if $\Regex$ is a stochastic regular expression, $\BRegex$ is the regular
expression obtained by removing probability annotations.

\begin{center}
  \begin{tabular}{l@{\ }l@{\ }l@{\ }l}
    % REGEX
    \Lens{} & \GEq{} & & $\IdentityLensOf{\BRegex}$\\
            & & \GBar{} & $\DisconnectOf{\BRegex}{\BRegexAlt}{\String}{\StringAlt}$ \\
            & & \GBar{} & $\IterateLensOf{\Lens}$ \\
            & & \GBar{} & $\ConcatLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\SwapLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\OrLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\ComposeLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\MergeROf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\MergeLOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\InvertOf{\Lens}$ \\
  \end{tabular}
\end{center}

@@

\[
  \centering
  \inferrule*
  {
  }
  {
    \IdentityLensOf{\BRegex} \OfType \BRegex \Leftrightarrow \BRegex
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App s$ & = & $s$\\
    $\CreateL{} \App s$ & = & $s$\\
    $\PutR{} \App s_1 \App s_2$ & = & $s_1$\\
    $\PutL{} \App s_1 \App s_2$ & = & $s_1$
  \end{tabular}
\end{center}

@@

\[
  \centering
  \inferrule*
  {
    \String \in \LanguageOf{\BRegex}\\
    \StringAlt \in \LanguageOf{\BRegexAlt}
  }
  {
    \DisconnectOf{\BRegex}{\BRegexAlt}{\String}{\StringAlt}
    \OfType \BRegex \Leftrightarrow \BRegexAlt
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String'$ & = & $\StringAlt$\\
    $\CreateL{} \App \StringAlt'''''$ & = & $\String$\\
    $\PutR{} \App \String''' \App \StringAlt'$ & = & $\StringAlt'$\\
    $\PutL{} \App \StringAlt' \App \String'$ & = & $\String'$
  \end{tabular}
\end{center}
Just as the identity lens ignores the second argument in puts, disconnect lenses
ignore the first in both puts and creates.  The data is unsynchronized in these
two formats, information from one format doesn't impact the other.

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
  }
  {
    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
    \Leftrightarrow
    \BRegexAlt_1 \Concat \BRegexAlt_2
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1 \String_2$ & = & $(\Lens_1.\CreateROf{\String_1})(\Lens_2.\CreateROf{\String_2})$\\
%    $\CreateL{} \App \StringAlt_1 \StringAlt_2$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
%    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_1\StringAlt_2)$ & = & $(\Lens_1.\PutROf{\String_1}{\StringAlt_1})(\Lens_2.\PutROf{\String_2}{\StringAlt_2})$\\
%    $\PutL{} \App (\StringAlt_1\StringAlt_2) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
%  \end{tabular}
%\end{center}

@@

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
  }
  {
    \SwapLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
    \Leftrightarrow
    \BRegexAlt_2 \Concat \BRegexAlt_1
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1\String_2$ & = & $(\Lens_2.\CreateROf{\String_2})(\Lens_1.\CreateROf{\String_1})$\\
%    $\CreateL{} \App \StringAlt_2\StringAlt_1$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
%    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_2\StringAlt_1)$ & = & $(\Lens_2.\PutROf{\String_2}{\StringAlt_2})(\Lens_1.\PutROf{\String_1}{\StringAlt_1})$\\
%    $\PutL{} \App (\StringAlt_2\StringAlt_1) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
%  \end{tabular}
%\end{center}
The swap combinator is similar to concat, though the second regular expression
is swapped.

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
  }
  {
    \OrLensOf{\Lens_1}{\Lens_2} \OfType
    \RegexOr{\BRegex_1}{\BRegex_2}
    \Leftrightarrow
    \RegexOr{\BRegexAlt_1}{\BRegexAlt_2}
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String$
    & =
    & $\begin{cases*}
      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2}$
      \end{cases*}$\\

    $\CreateL{} \App \StringAlt$
    & =
    & $\begin{cases*}
      \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
      \Lens_2.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
      \end{cases*}$\\

    $\PutR{} \App \String \App \StringAlt$
    & =
    & $\begin{cases*}
        \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_1} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
        \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_2}$\\
        \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_2}$\\
        \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$
      \end{cases*}$\\

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$
      \end{cases*}$\\
  \end{tabular}
\end{center}
The \OrLens lens deals with data that can come in one form or another. If the
data gets changed from one format to the other, information in the old format is
lost.

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
  }
  {
    \IterateLensOf{\Lens} \OfType
    \StarOf{\BRegex}
    \Leftrightarrow
    \StarOf{\BRegexAlt}
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1\ldots\String_n$
%    & =
%    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
%    
%    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
%    & =
%    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
%    
%    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
%    & =
%    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
%        \Lens.\CreateROf{\String_i} & otherwise
%      \end{cases*}$\\
%    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
%    & =
%    & $\String_1'\ldots\String_n'$ where $\String_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
%        \Lens.\CreateROf{\StringAlt_i} & otherwise
%      \end{cases*}$
%  \end{tabular}
%\end{center}
The \IterateLens lens deals with iterated data.

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt\\
    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt
  }
  {
    \MergeROf{\Lens_1}{\Lens_2} \OfType
    \RegexOr{\BRegex_1}{\BRegex_2}
    \Leftrightarrow
    \BRegexAlt
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String$
    & =
    & $\begin{cases*}
      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2}$
      \end{cases*}$\\
